Binning Strategies and Related Loss for Large Data

### 2015-04-24 Paper.Rnw
Remove old plot comparing standard and random, now refered to new figure.
Plot caption improvements

The standard rectangular binning function $b(.) : (x_i,y_i) \rightarrow (x^\ast_\ell,y^\ast_\ell)$ is defined as 
%
\begin{eqnarray}\label{standrecbin1}
b(x_i,y_i) = (b_X(x_i),b_Y(y_i))
\end{eqnarray}
%
where $b_X(x_i)$ and $b_Y(y_i)$ are the univariate standard binning algorithms for the X and Y dimensions respectively. The random rectangular binning function, $b^r(\cdot): (x_i,y_i) \rightarrow (x^\ast_\ell,y^\ast_\ell)$ can similarly be defined as
%
\begin{eqnarray}\label{randrecbin1}
b^r(x_i,y_i) = (b^r_X(x_i),b^r_Y(y_i))
\end{eqnarray}
%
where $b^r_X(x_i)$ and $b^r_Y(y_i)$ are univariate random binning algorithms for each dimension. Figure~\ref{fig:BinningTutorial} provides an illustration of each binning process extended to a two dimensional situation.

<<binningDemo,echo=F,include=F,cache=T>>=
      library(mvtnorm)
      set.seed(12)
      d <- rmvnorm(50, c(10, 25), matrix(c(50, 15, 15, 50), nrow=2))
      d <- data.frame(d)
      names(d) <- c("x", "y")
      d <- subset(d, (x>-20 & x<40 & y>0 & y<50),drop=TRUE)
      d.rand <- d
      dnew <- binStd(cbind(d-5, Freq=1), c(10, 10))
      dnew$x <- dnew$x+5
      dnew$y <- dnew$y+5
      d$bin <- dnew$x%in%c(5,15) & dnew$y==25
      d$a <- .33 + .66*d$bin
      d.orig <- d

      d$xbin <- NA
      d$ybin <- NA
      idx <- which(dnew$x%in%c(5, 15) & dnew$y==25)
      d$xbin[idx] <- dnew$x[idx]
      d$ybin[idx] <- 25

temp <- data.frame(   
          x=c( -10,   0,  10,  20,  30,  -10, -10, -10, -10), 
       xend=c( -10,   0,  10,  20,  30,   30,  30,  30,  30), 
          y=c( 10,   10,  10,  10,  10,   10,  20,  30,  40), 
       yend=c( 40,   40,  40,  40,  40,   10,  20,  30,  40), 
      group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9))
      qplot(data=d[,], x=x, y=y, alpha=a,  geom="point", 
            xlim=c(-20, 40), ylim=c(0, 50)) +
        annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
                 fill="blue", alpha=.25) +
        annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), 
                 fill="green", alpha=.25) +
        scale_alpha(range=c(.5, 1), guide=FALSE) +
      #  scale_colour_identity(guide=FALSE) + 
      #  scale_shape(solid=T,guide=FALSE) +
        theme_bw() +
      #  annotate("point", x=rep(seq(-15, 35, 10),each=7), 
      #         y=rep(seq(-5, 55, 10), times=6), 
      #         shape='+', size=8, colour="gray90") + 
        geom_point(data=d[which(!is.na(d$xbin)),], 
                   aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), 
                   size=I(3), aes.inherit=FALSE) + 
        scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), 
                           labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
        scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", 
                            labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
        annotate("segment", x=temp$x, y=temp$y, 
                 xend=temp$xend, yend=temp$yend, group=temp$group) + 
        ylim(c(10, 40)) + xlim(c(-10,30))
      ggsave("./figure/Binning-DataAndBin.pdf", width=5, height=5, units="in")

      d.orig$bin <- dnew$x==15 & dnew$y==25
      d.orig$a <- .33+.66*d.orig$bin
      d.orig$l <- "Data outside bin"
      d.orig$l[d.orig$bin] <- "Binned Data"
      d.orig$size <- 2
      d.orig$size[d.orig$bin] <- 3
      d.orig <- rbind(d.orig, data.frame(x=15, y=25, a=1, l="Visual Center", size=2,bin=TRUE),
        data.frame(x=mean(d.orig$x[d.orig$bin]), y=mean(d.orig$y[d.orig$bin]),                                                                                    
        l="Numerical Center", a=1, size=2, bin=TRUE))
#       ggplot(data=d.orig, aes(x=x, y=y, shape=l, size=l, colour=l)) + 
#         annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
#                  fill="blue", alpha=.10) + 
#         geom_point() +
#         scale_shape_manual(name="", values=c("Data outside bin" = 16, 
#                                              "Binned Data" = 15, 
#                                              "Numerical Center" = 17, 
#                                              "Visual Center" = 18)) +
#         scale_colour_manual(name="", values=c("Data outside bin" = "grey60", 
#                                               "Binned Data" = "blue",
#                                               "Numerical Center" = "blue4", 
#                                               "Visual Center" = "blue4"))+
#         scale_size_manual(name="", values=c("Data outside bin" = 3, "Binned Data"=3,
#                                             "Numerical Center" = 5, 
#                                             "Visual Center" = 5))+
#         theme_bw() +
#         annotate("point", x=15, y=25, shape=18, size=5, colour="blue4") + 
#         annotate("point", x=d.orig$x[which(d.orig$l=="Numerical Center")], 
#                  y=d.orig$y[which(d.orig$l=="Numerical Center")], shape=17, size=5, 
#                  colour="blue4") +
# #        annotate("segment", x=temp$x, y=temp$y, xend=temp$xend, yend=temp$yend, group=temp$group) + 
# #         annotate("text", x=d$x[which(d$l=="Numerical Center")], 
# #                  y=d$y[which(d$l=="Numerical Center")], 
# #                  label=expression(paste('(', over(phantom(0),x)[i], ',', over(phantom(0),y)[i],')'))) +
#         scale_x_continuous(breaks=c(10, 15, 20), limits=c(8, 22), labels=c(expression(x^{"*"}-b[x]/2), expression(x^{"*"}), expression(x^{"*"}+b[x]/2)), name='') + 
#         scale_y_continuous(breaks=c(20, 25, 30), limits=c(18, 32), labels=c(expression(y^{"*"}-b[y]/2), expression(y^{"*"}), expression(y^{"*"}+b[y]/2)), name='')
#       ggsave("./figure/Binning-DataVisualNumericalCenter.pdf", width=5, height=4, units="in")
@

<<RandomBinDemo,echo=F,include=FALSE>>=
nl <- ncol(d.rand)
dnew <- d.rand
binning <- c(10, 10)
for(x in 1:(nl)){
  y <- d.rand[,x] - binning[x]/2
dnew[,x] <- binning[x]*(floor(y/binning[x]) + 
  sapply((y%%binning[x])/binning[x], function(p) rbinom(1, 1, p))) + binning[x]/2
}
temp <- data.frame(   
          x=c( -10,   0,  10,  20,  30,  -10, -10, -10, -10), 
       xend=c( -10,   0,  10,  20,  30,   30,  30,  30,  30), 
          y=c( 10,   10,  10,  10,  10,   10,  20,  30,  40), 
       yend=c( 40,   40,  40,  40,  40,   10,  20,  30,  40), 
      group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9))

dnew$id <- interaction(dnew[,1:(nl)])
idx <- which(dnew$x%in%c(5, 15) & dnew$y==25)
d.rand$bin <- dnew$x%in%c(5, 15) & dnew$y==25
d.rand$a <- .33 + .66*d.rand$bin
d.rand$xbin <- NA
d.rand$ybin <- NA
d.rand$xbin[idx] <- dnew$x[idx]
d.rand$ybin[idx] <- 25
# res <- ddply(dnew, .(id), summarize, n=length(id), fsum=sum(Freq))
qplot(data=d.rand[,], x=x, y=y, alpha=a,  geom="point", xlim=c(-20, 40), ylim=c(0, 50)) +
  annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), fill="blue", alpha=.25) +
  annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), fill="green", alpha=.25) +
  scale_alpha(range=c(.5, 1), guide=FALSE) +
  scale_colour_identity(guide=FALSE) + 
  scale_shape(solid=T,guide=FALSE) +
  theme_bw() +
#  annotate("point", x=rep(seq(-15, 35, 10),each=7), 
#        y=rep(seq(-5, 55, 10), times=6), 
#         shape='+', size=8, colour="gray90") + 
  geom_point(data=d.rand[which(!is.na(d.rand$xbin)),], 
             aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), size=I(3), aes.inherit=FALSE) + 
  scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
  scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
  annotate("segment", x=temp$x, y=temp$y, xend=temp$xend, yend=temp$yend, group=temp$group) + ylim(c(10, 40)) + xlim(c(-10,30))
ggsave("./figure/RandomBinningDataPoints2.pdf",width=5, height=5)
@

\begin{figure}[h]
  \centering
  \subfloat[Standard Binning]{\includegraphics[keepaspectratio=true, width=2.5in]{./figure/Binning-DataAndBin.pdf}}
%   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
%   {./figure/RandomBinningDataPoints1}}
  \subfloat[Random Binning]{\includegraphics[keepaspectratio=TRUE, width=2.5in]
  {./figure/RandomBinningDataPoints2.pdf}}
  \caption{Visualization of how data pairs are assigned to bins centered at (5, 25) and (15,25) under the two rectangular binning algorithms. Points in the standard algorithm are contained within the defined bin region, where points from the random algorithm may be assigned to any of the surrounding bin centers.}
  \label{fig:BinningTutorial}
\end{figure}

@@

% recover some of the numerical information lost due to binning by storing the numerical center $(\bar{x},\bar{y})$ instead of the visual center $(x^\ast,y^\ast)$ for use in subsequent calculations. Storing the binned data as the arithmetic mean rather than the visual center reduces the total loss, and the visual center can be recovered from the bin width and  bin center $(O_X, O_Y)$, see table \ref{RecoveryComparison} for a numeric example for this.

@@

%Information recovery comes with a two-fold cost: the additional storage of numeric bin centers triples the amount of information stored, and the computational cost to compute numeric bin centers is O($nm^2$), where $n$ is the number of data points and $m^2$, the total number of bins. O($nm^2$) $\approx$ O($n$), as $m$ is generally negligible compared to $n$.

% In spite of the costs, information recovery is still useful for very large data, where the unbinned data set is  computationally intractable. Minimizing numerical loss in this manner provides for more accurate calculations from the data, while maintaining the storage space advantages of binning.

@@

<<UniformStripes, echo=F,include=T,eval=T,fig.width=8, fig.height=5, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Binned scatterplots of coarse uniform data with square bins of size 1, 4 and 5">>= 
# Commented out code for data generation: run once then recomment
### Create data to demonstrate stripes from improper binwidths
# binoutcoarse1 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,1,1,type="standard")[[1]]
# binoutcoarse4 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,4,4,type="standard")[[1]]
# binoutcoarse5 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,5,5,type="standard")[[1]]
# binoutcoarse1Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,1,1,type="random")[[1]]
# binoutcoarse4Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,4,4,type="random")[[1]]
# binoutcoarse5Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,5,5,type="random")[[1]]
# 
# binoutcoarse1$BinSize <- "1 X 1 Unit" ; binoutcoarse1$BinningType <- "Standard" 
# binoutcoarse4$BinSize <- "4 X 4 Unit" ; binoutcoarse4$BinningType <- "Standard" 
# binoutcoarse5$BinSize <- "5 X 5 Unit" ; binoutcoarse5$BinningType <- "Standard" 
# binoutcoarse1Rand$BinSize <- "1 X 1 Unit" ; binoutcoarse1Rand$BinningType <- "Random" 
# binoutcoarse4Rand$BinSize <- "4 X 4 Unit" ; binoutcoarse4Rand$BinningType <- "Random" 
# binoutcoarse5Rand$BinSize <- "5 X 5 Unit" ; binoutcoarse5Rand$BinningType <- "Random" 
# 
# allbindat <- rbind(binoutcoarse1,binoutcoarse4,binoutcoarse5,
#                    binoutcoarse1Rand,binoutcoarse4Rand,binoutcoarse5Rand)
# write.csv(allbindat,"StripeBinData.csv", row.names=FALSE)

@@

<<ExponentialOriginPlots, echo=F,include=T,eval=T,fig.width=8, fig.height=3, out.width='.9\\linewidth', fig.pos='H',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Binned scatterplots for the fine exponential data using standard binning with 10 by 10 unit bins with origins at (0,0) and (-9,-9). These binnings have net spatial losses of 391067 units and 420458 units, respectively">>=

>>>>

The standard rectangular binning function $b(.) : (x_i,y_i) \rightarrow (x^\ast_\ell,y^\ast_\ell)$ is defined as 
%
\begin{eqnarray}\label{standrecbin1}
b(x_i,y_i) = (b_X(x_i),b_Y(y_i))
\end{eqnarray}
%
where $b_X(x_i)$ and $b_Y(y_i)$ are the univariate standard binning algorithms for the X and Y dimensions respectively. The random rectangular binning function, $b^r(\cdot): (x_i,y_i) \rightarrow (x^\ast_\ell,y^\ast_\ell)$ is similarly be defined as
%
\begin{eqnarray}\label{randrecbin1}
b^r(x_i,y_i) = (b^r_X(x_i),b^r_Y(y_i))
\end{eqnarray}
%
where $b^r_X(x_i)$ and $b^r_Y(y_i)$ are univariate random binning algorithms for each dimension. Figure~\ref{fig:spatlossdemo2} provides an illustration of each binning process extended to a two dimensional situation.
% <<binningDemo,echo=F,include=F,cache=T>>=
%       library(mvtnorm)
%       set.seed(12)
%       d <- rmvnorm(50, c(10, 25), matrix(c(50, 15, 15, 50), nrow=2))
%       d <- data.frame(d)
%       names(d) <- c("x", "y")
%       d <- subset(d, (x>-20 & x<40 & y>0 & y<50),drop=TRUE)
%       d.rand <- d
%       dnew <- binStd(cbind(d-5, Freq=1), c(10, 10))
%       dnew$x <- dnew$x+5
%       dnew$y <- dnew$y+5
%       d$bin <- dnew$x%in%c(5,15) & dnew$y==25
%       d$a <- .33 + .66*d$bin
%       d.orig <- d
%       d$xbin <- NA
%       d$ybin <- NA
%       idx <- which(dnew$x%in%c(5, 15) & dnew$y==25)
%       d$xbin[idx] <- dnew$x[idx]
%       d$ybin[idx] <- 25
% temp <- data.frame(   
%           x=c( -10,   0,  10,  20,  30,  -10, -10, -10, -10), 
%        xend=c( -10,   0,  10,  20,  30,   30,  30,  30,  30), 
%           y=c( 10,   10,  10,  10,  10,   10,  20,  30,  40), 
%        yend=c( 40,   40,  40,  40,  40,   10,  20,  30,  40), 
%       group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9))
%       qplot(data=d[,], x=x, y=y, alpha=a,  geom="point", 
%             xlim=c(-20, 40), ylim=c(0, 50)) +
%         annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
%                  fill="blue", alpha=.25) +
%         annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), 
%                  fill="green", alpha=.25) +
%         scale_alpha(range=c(.5, 1), guide=FALSE) +
%       #  scale_colour_identity(guide=FALSE) + 
%       #  scale_shape(solid=T,guide=FALSE) +
%         theme_bw() +
%       #  annotate("point", x=rep(seq(-15, 35, 10),each=7), 
%       #         y=rep(seq(-5, 55, 10), times=6), 
%       #         shape='+', size=8, colour="gray90") + 
%         geom_point(data=d[which(!is.na(d$xbin)),], 
%                    aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), 
%                    size=I(3), aes.inherit=FALSE) + 
%         scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), 
%                            labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
%         scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", 
%                             labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
%         annotate("segment", x=temp$x, y=temp$y, 
%                  xend=temp$xend, yend=temp$yend, group=temp$group) + 
%         ylim(c(10, 40)) + xlim(c(-10,30))
%       ggsave("./figure/Binning-DataAndBin.pdf", width=5, height=5, units="in")
%       d.orig$bin <- dnew$x==15 & dnew$y==25
%       d.orig$a <- .33+.66*d.orig$bin
%       d.orig$l <- "Data outside bin"
%       d.orig$l[d.orig$bin] <- "Binned Data"
%       d.orig$size <- 2
%       d.orig$size[d.orig$bin] <- 3
%       d.orig <- rbind(d.orig, data.frame(x=15, y=25, a=1, l="Visual Center", size=2,bin=TRUE),
%         data.frame(x=mean(d.orig$x[d.orig$bin]), y=mean(d.orig$y[d.orig$bin]),                                                                                    
%         l="Numerical Center", a=1, size=2, bin=TRUE))
% #       ggplot(data=d.orig, aes(x=x, y=y, shape=l, size=l, colour=l)) + 
% #         annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
% #                  fill="blue", alpha=.10) + 
% #         geom_point() +
% #         scale_shape_manual(name="", values=c("Data outside bin" = 16, 
% #                                              "Binned Data" = 15, 
% #                                              "Numerical Center" = 17, 
% #                                              "Visual Center" = 18)) +
% #         scale_colour_manual(name="", values=c("Data outside bin" = "grey60", 
% #                                               "Binned Data" = "blue",
% #                                               "Numerical Center" = "blue4", 
% #                                               "Visual Center" = "blue4"))+
% #         scale_size_manual(name="", values=c("Data outside bin" = 3, "Binned Data"=3,
% #                                             "Numerical Center" = 5, 
% #                                             "Visual Center" = 5))+
% #         theme_bw() +
% #         annotate("point", x=15, y=25, shape=18, size=5, colour="blue4") + 
% #         annotate("point", x=d.orig$x[which(d.orig$l=="Numerical Center")], 
% #                  y=d.orig$y[which(d.orig$l=="Numerical Center")], shape=17, size=5, 
% #                  colour="blue4") +
% # #        annotate("segment", x=temp$x, y=temp$y, xend=temp$xend, yend=temp$yend, group=temp$group) + 
% # #         annotate("text", x=d$x[which(d$l=="Numerical Center")], 
% # #                  y=d$y[which(d$l=="Numerical Center")], 
% # #                  label=expression(paste('(', over(phantom(0),x)[i], ',', over(phantom(0),y)[i],')'))) +
% #         scale_x_continuous(breaks=c(10, 15, 20), limits=c(8, 22), labels=c(expression(x^{"*"}-b[x]/2), expression(x^{"*"}), expression(x^{"*"}+b[x]/2)), name='') + 
% #         scale_y_continuous(breaks=c(20, 25, 30), limits=c(18, 32), labels=c(expression(y^{"*"}-b[y]/2), expression(y^{"*"}), expression(y^{"*"}+b[y]/2)), name='')
% #       ggsave("./figure/Binning-DataVisualNumericalCenter.pdf", width=5, height=4, units="in")
% <<RandomBinDemo,echo=F,include=FALSE>>=
% nl <- ncol(d.rand)
% dnew <- d.rand
% binning <- c(10, 10)
% for(x in 1:(nl)){
%   y <- d.rand[,x] - binning[x]/2
% dnew[,x] <- binning[x]*(floor(y/binning[x]) + 
%   sapply((y%%binning[x])/binning[x], function(p) rbinom(1, 1, p))) + binning[x]/2
% temp <- data.frame(   
%           x=c( -10,   0,  10,  20,  30,  -10, -10, -10, -10), 
%        xend=c( -10,   0,  10,  20,  30,   30,  30,  30,  30), 
%           y=c( 10,   10,  10,  10,  10,   10,  20,  30,  40), 
%        yend=c( 40,   40,  40,  40,  40,   10,  20,  30,  40), 
%       group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9))
% dnew$id <- interaction(dnew[,1:(nl)])
% idx <- which(dnew$x%in%c(5, 15) & dnew$y==25)
% d.rand$bin <- dnew$x%in%c(5, 15) & dnew$y==25
% d.rand$a <- .33 + .66*d.rand$bin
% d.rand$xbin <- NA
% d.rand$ybin <- NA
% d.rand$xbin[idx] <- dnew$x[idx]
% d.rand$ybin[idx] <- 25
% # res <- ddply(dnew, .(id), summarize, n=length(id), fsum=sum(Freq))
% qplot(data=d.rand[,], x=x, y=y, alpha=a,  geom="point", xlim=c(-20, 40), ylim=c(0, 50)) +
%   annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), fill="blue", alpha=.25) +
%   annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), fill="green", alpha=.25) +
%   scale_alpha(range=c(.5, 1), guide=FALSE) +
%   scale_colour_identity(guide=FALSE) + 
%   scale_shape(solid=T,guide=FALSE) +
%   theme_bw() +
% #  annotate("point", x=rep(seq(-15, 35, 10),each=7), 
% #        y=rep(seq(-5, 55, 10), times=6), 
% #         shape='+', size=8, colour="gray90") + 
%   geom_point(data=d.rand[which(!is.na(d.rand$xbin)),], 
%              aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), size=I(3), aes.inherit=FALSE) + 
%   scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
%   scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
%   annotate("segment", x=temp$x, y=temp$y, xend=temp$xend, yend=temp$yend, group=temp$group) + ylim(c(10, 40)) + xlim(c(-10,30))
% ggsave("./figure/RandomBinningDataPoints2.pdf",width=5, height=5)
% \begin{figure}[h]
%   \subfloat[Standard Binning]{\includegraphics[keepaspectratio=true, width=2.5in]{./figure/Binning-DataAndBin.pdf}}
% %   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
% %   {./figure/RandomBinningDataPoints1}}
%   \subfloat[Random Binning]{\includegraphics[keepaspectratio=TRUE, width=2.5in]
%   {./figure/RandomBinningDataPoints2.pdf}}
%   \caption{Visualization of how data pairs are assigned to bins centered at (5, 25) and (15,25) under the two rectangular binning algorithms. Points in the standard algorithm are contained within the defined bin region, where points from the random algorithm may be assigned to any of the surrounding bin centers.}
%   \label{fig:BinningTutorial}

@@

@@

@@

<<UniformStripes, echo=F,include=T,eval=T,fig.width=8, fig.height=5, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Binned scatterplots of coarse uniform data with 1X1, 4X4 and 5X5 square bins. For standard binning, the 1X1 bins leaves white-space gaps between bins, and 5X5 bins cause artificial density stripes; whereas random binning smoothes density over these poor choices in bin dimensions.">>= 
# Commented out code for data generation: run once then recomment
### Create data to demonstrate stripes from improper binwidths
# binoutcoarse1 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,1,1,type="standard")[[1]]
# binoutcoarse4 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,4,4,type="standard")[[1]]
# binoutcoarse5 <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,5,5,type="standard")[[1]]
# binoutcoarse1Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,1,1,type="random")[[1]]
# binoutcoarse4Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,4,4,type="random")[[1]]
# binoutcoarse5Rand <- RectBin2d(datlist[[1]]$roundx,datlist[[1]]$roundy,0,0,5,5,type="random")[[1]]
# 
# binoutcoarse1$BinSize <- "1 X 1 Unit" ; binoutcoarse1$BinningType <- "Standard" 
# binoutcoarse4$BinSize <- "4 X 4 Unit" ; binoutcoarse4$BinningType <- "Standard" 
# binoutcoarse5$BinSize <- "5 X 5 Unit" ; binoutcoarse5$BinningType <- "Standard" 
# binoutcoarse1Rand$BinSize <- "1 X 1 Unit" ; binoutcoarse1Rand$BinningType <- "Random" 
# binoutcoarse4Rand$BinSize <- "4 X 4 Unit" ; binoutcoarse4Rand$BinningType <- "Random" 
# binoutcoarse5Rand$BinSize <- "5 X 5 Unit" ; binoutcoarse5Rand$BinningType <- "Random" 
# 
# allbindat <- rbind(binoutcoarse1,binoutcoarse4,binoutcoarse5,
#                    binoutcoarse1Rand,binoutcoarse4Rand,binoutcoarse5Rand)
# write.csv(allbindat,"StripeBinData.csv", row.names=FALSE)

@@

<<ExponentialOriginPlots, echo=F,include=T,eval=T,fig.width=8, fig.height=3, out.width='.9\\linewidth', fig.pos='H',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Binned scatterplots for the fine exponential data using standard binning with 10 by 10 unit bins with origins at (0,0) and (-9,-9). The binning origin at (0,0) leads to 9\\% less net spatial loss than (-9,-9); with losses of 391067 units and 420458 units, respectively.">>=
